---
title: "pm566-midterm-draft2"
author: "Misha Khan"
date: "2022-10-22"
output: github_document
always_allow_html: true
---

```{r}
library(stringr)
library(dplyr)
library(tidyverse)
library(rvest)
library(xml2)
library(dbplyr)
library(dplyr)
library(skimr) 
library(ggplot2)
library(data.table)
library(lubridate)
library(leaflet)
library(webshot)
```

Topic: Crime Data from 2020 to Present in Los Angeles
Missing data (0, 0) for long and lat are for privacy
URL: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8
```{r}
crimeraw <- read.csv("Crime_Data_from_2020_to_Present.csv")
```

Data Cleaning
Rename cols to make it more meaningful
```{r}
colnames(crimeraw) <- c('Record No',
                        'Date Reported',
                        'Date Occurred',
                        'Time Occurred',
                        'Area No',
                        'Area',
                        'Rpt Dist No', #drop
                        'Part', #drop
                        'Crime Code',
                        'Crime',
                        'MO', #drop
                        'Victim Age',
                        'Victim Sex',
                        'Victim Ethnicity',
                        'Premise Code',
                        'Premise',
                        'Weapon Code',
                        'Weapon',
                        'Case Code',
                        'Case Status',
                        'Crime Code 1',
                        'Crime Code 2',
                        'Crime Code 3',
                        'Crime Code 4',
                        'Location',
                        'Cross Street',
                        'Lat',
                        'Lon'
                        )

crimeraw <- select(crimeraw, -c('Rpt Dist No','Part', 'MO','Crime Code 1', 'Crime Code 2', 'Crime Code 3', 'Crime Code 4', 'Cross Street'))
```

Reformat date variables (EDA: Date when most crimes occur)
```{r}
crimeraw$`Date Reported` <- as.Date(crimeraw$`Date Reported`, format = "%m/%d/%Y")
crimeraw$`Date Occurred` <- as.Date(crimeraw$`Date Occurred`, format = "%m/%d/%Y")
```

Change miltary to standard time (Possible EDA: Time when most crime occur)
```{r}
crimeraw$`Time Occurred` <- format(strptime(substr(as.POSIXct(sprintf("%04.0f", crimeraw$`Time Occurred`), 
                                 format="%H%M"), 12, 16), '%H:%M'), '%I:%M %p')
```

View missing data
```{r}
skimr::skim(crimeraw)
```

MISSING: Weapon Code
EMPTY: Victim Sex, Victim Ethnicity, Premise, Weapon
IF missing, change to 999
FIX: empty cells will not change to value or NA
```{r}
crimeraw$`Weapon Code` <- crimeraw$`Weapon Code` %>% replace_na(999)
```


New dataset created
*START HERE*
```{r}
#Save as changes as a new file
#write_rds(crimeraw, "crimedatnew.rds")
crimedat <- read_rds("crimedatnew.rds")
crimecopy <- crimedat
crimedat <- as.data.frame(crimedat)
```

EDA: What is the amount of crime per area in LA from 2020-2022?
What is the proportion of crime?
```{r}
crime_frequency <-count(crimedat, Area)
nc <- sum(crime_frequency$n)

crime_frequency %>% 
  arrange(desc(n)) %>%
  mutate(Percent = round( n / nc * 100, 2))  %>%
knitr::kable()
```
```{r}
crime_frequency <- crime_frequency %>% arrange(desc(n))
```


```{r}
# Pie Chart with Percentages
slices <- crime_frequency$n[1:10]
lbls <- c("Central", "77th Street", "Pacific", "Southwest", "Hollywood", "Southeast", "N Hollywood", "Newton", "Olympic", "Wilshire")
pct <- round(slices/sum(slices)*100, 2)
lbls <- paste(lbls, pct) # add percents to labels
lbls <- paste(lbls,"%",sep="") # ad % to labels
pie(slices,labels = lbls, col=rainbow(length(lbls)),
   main="Proportion of Top 10 Areas in Crime")
```
Visualization of top 10 areas
```{r}
v1 <- table(crimedat$`Area`)
barplot(sort(v1, T)[1:10], las = 2, col = rainbow(12), cex.names= .7, main = "Top 10 Areas with Crime", xlab = "Area", ylab = "Count")     
```

Get month before counting crimes by month and area.

```{r}
crimedat$Month = format(crimedat$`Date Occurred`, "%B")
```

#MY EDITS
```{r}
#Create CrimeCount col for the Area
crime_count_month <-   #Instead of changing org dataframe (crimedat), I create a new df
  crimedat %>%
  group_by(Month, `Area`) %>% 
  summarize(CrimeCount = n())
```


Find month with the highest crime (3 years of data)
```{r}
crime_count_month %>%
ggplot(aes(x = Month, y = CrimeCount, fill = Area)) +
  geom_bar(stat = "identity")
```


```{r}
crimedat$Year = format(crimedat$`Date Occurred`, "%Y")
```

```{r}
#Create CrimeCount col for the Area
crime_count_year <-   
  crimedat %>%
  group_by(Year, `Area`) %>% 
  summarize(CrimeCount = n())
```


Find year with the highest crime (3 years of data)
```{r}
crime_count_year %>%
ggplot(aes(x = Year, y = CrimeCount, fill = Area)) +
  geom_bar(stat = "identity")
```

Find the most reported crime
```{r}
crimedat %>%
  group_by(`Crime Code`, `Crime`) %>%
  summarise(n = n())  %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```
Create a most common crimes df
```{r}
crime_top_areas <- crimedat %>%
  select(`Crime Code`, `Crime`, `Year`, `Lat`, `Lon`)  
```
Subset to top 10 using Crime Code in 2021
```{r}
crime_top_areas <- subset(crime_top_areas, 
                          `Crime Code` %in% c("510", "624", "330", "740", "310", "230", "354", "440", "626", "420") &
                          `Year` %in% c("2021"))
```

Map color palette
```{r}
crime.pal <- colorFactor(c('red', 'yellow', 'blue', 'green'), domain = crime_top_areas$`Crime`)
```

Map of the area of top 10 crimes in 2021
```{r}
crime_map <- 
  leaflet(crime_top_areas) %>%  
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    lat = ~Lat, 
    lng = ~Lon,
    label = ~paste0(crime_top_areas$`Crime`),
    color = ~ crime.pal(crime_top_areas$`Crime`),
    opacity = 0.5,
    fillOpacity = 1,
    radius = 500
    ) %>%
  addLegend('bottomleft', 
            pal = crime.pal,  
            values = crime_top_areas$`Crime`,
            title = 'Top Crimes per Area', 
            opacity = 1)
crime_map
```

Finding the outlier coordinate near Africa
```{r}
range(crime_top_areas$Lat)
range(crime_top_areas$Lon)
```

```{r}
library(CoordinateCleaner)
#cc_outl(crime_top_areas,
#        method = "quantile", 
#        value = "flagged")
```